Time slice analyses

### 2017-11-07 analyses/02-time-slice-analyses.Rmd
pipelined analysis (test on test data not real!) and plots

@@

## Ages
out1 <- run.all.disparity.strat(morphospace, tree, type = "Age", FADLAD, inc.nodes = TRUE,
                                bootstraps = 100, metric = c(sum, variances), metric.name = "sum_var")
readr::write_csv(out1, path = paste0("../outputs/age", "_", slug, ".csv"))

## Epochs
out2 <- run.all.disparity.strat(morphospace, tree, type = "Epoch", FADLAD, inc.nodes = TRUE,
                        bootstraps = 100, metric = c(sum, variances), metric.name = "sum_var")
readr::write_csv(out2, path = paste0("../outputs/epoch", "_", slug, ".csv"))
```

@@

@@

Plots
Statistics

>>>>

> TG: for the following pipeline, I've reduced the number of rarefaction that highly increased the calculation time.
We now only calculate the disparity for the non-rarefied subsamples and the subsamples rarefied to contain only 3 taxa (the lowest level of rarefaction).
We can use the full rarefaction if we want to plot rarefaction curves later on.

```{r}
## Testing data: TO REMOVE
#TG: it's just faster this way
#TG: also, I've only tested it on "Epoch" since it take ages to run, should be pipelined all right though
data(BeckLee_mat99)
data(BeckLee_tree)
data(BeckLee_ages)

morphospace <- BeckLee_mat99
tree <- BeckLee_tree
FADLAD <- BeckLee_ages

@@

## Ages (rarefied to 3 taxa only)
out1 <- run.all.disparity.wrapper(morphospace, tree, type = "Age", FADLAD, inc.nodes = TRUE,
                                bootstraps = 100, metric = c(sum, variances),
                                metric.name = "sum_var", rarefaction = 3)
#readr::write_csv(out1, path = paste0("../outputs/age", "_", slug, ".csv"))

## Epochs (rarefied to 3 taxa only)
out2 <- run.all.disparity.wrapper(morphospace, tree, type = "Epoch", FADLAD, inc.nodes = TRUE,
                        bootstraps = 100, metric = c(sum, variances), metric.name = "sum_var",
                        rarefaction = FALSE)
#readr::write_csv(out2, path = paste0("../outputs/epoch", "_", slug, ".csv"))
```

@@

```
<!-- Need to think more about the outputs here, currently they don't make much sense as there are no column headers -->

@@

The results as a table

```{r, eval = FALSE}
knitr::kable(out2$results$stratigraphy, caption = "Disparity per stratigraphic time: unequal time bins and non-equidistant time slices")
knitr::kable(out2$results$duration, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the median stratigraphic epoch duration")
knitr::kable(out2$results$number, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the number of stratigraphic epochs")
```

This is not really digest and we should find a better way to present that.
Anyway, these boring tables are always good for the supplementaries.

Plotting wise, we have to solutions, either using `dispRity`

```{r, eval = FALSE}
op <- par(bty = "n")
colors <- palette() #TG: that can be made fancier

## STRATIGRAPHY
## Plotting the continuous data
plot(out2$object$stratigraphy[[2]], quantile = 1, col = colors[1], ylim = c(1,4), main = "Equal duration and stratigraphys equal to stratigraphy", xlab = "sum of ranges")
plot(out2$object$stratigraphy[[3]], quantile = 1, col = colors[2], ylim = c(1,4), add = TRUE)
plot(out2$object$stratigraphy[[4]], quantile = 1, col = colors[3], ylim = c(1,4), add = TRUE)
plot(out2$object$stratigraphy[[5]], quantile = 1, col = colors[4], ylim = c(1,4), add = TRUE)
## Plotting the discrete data
plot(out2$object$stratigraphy[[1]], quantile = 1, type = "c", col = colors[5], ylim = c(1,4), add = TRUE, density = 0)
#TG: density = 0 is to not show the huge confidence interval, even for quantile = 1!
legend(x = "bottomright", legend = c("acctran", "deltran", "punctuated", "gradual", "time bins"), col = colors[1:5], lty = 1)
```

This on is interesting with no time bins after around 39 Mya and a drastic drop!

```{r, eval = FALSE}
op <- par(bty = "n")
colors <- palette() #TG: that can be made fancier
## DURATION
## Plotting the continuous data
plot(out2$object$duration[[2]], quantile = 1, col = colors[1], ylim = c(1,4), main = "Equal duration and durations equal to stratigraphy", xlab = "sum of ranges")
plot(out2$object$duration[[3]], quantile = 1, col = colors[2], ylim = c(1,4), add = TRUE)
plot(out2$object$duration[[4]], quantile = 1, col = colors[3], ylim = c(1,4), add = TRUE)
plot(out2$object$duration[[5]], quantile = 1, col = colors[4], ylim = c(1,4), add = TRUE)
## Plotting the discrete data
plot(out2$object$duration[[1]], quantile = 1, type = "c", col = colors[5], ylim = c(1,4), add = TRUE, density = 0)
legend(x = "bottomright", legend = c("acctran", "deltran", "punctuated", "gradual", "time bins"), col = colors[1:5], lty = 1)
```

Same absence of data after 39 but no drop!

```{r, eval = FALSE}
op <- par(bty = "n")
colors <- palette() #TG: that can be made fancier
## NUMBERS
## Plotting the continuous data
plot(out2$object$number[[2]], quantile = 1, col = colors[1], ylim = c(1,4), main = "Equal duration and numbers equal to stratigraphy", xlab = "sum of ranges")
plot(out2$object$number[[3]], quantile = 1, col = colors[2], ylim = c(1,4), add = TRUE)
plot(out2$object$number[[4]], quantile = 1, col = colors[3], ylim = c(1,4), add = TRUE)
plot(out2$object$number[[5]], quantile = 1, col = colors[4], ylim = c(1,4), add = TRUE)
## Plotting the discrete data
plot(out2$object$number[[1]], quantile = 1, type = "c", col = colors[5], ylim = c(1,4), add = TRUE)
legend(x = "bottomright", legend = c("acctran", "deltran", "punctuated", "gradual", "time bins"), col = colors[1:5], lty = 1)

This one is fairly similar but starts earlier in time.

Another way would be to do histograms the same way we drew it when we designed the project but I'll work on that tomorrow I guess.
